---
title: "Machine Learning Project"
author: "Carole U"
date: "14 May 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

This document is a report on the Weight Lifting exercises dataset as part of Coursera's Data Science machine learning class. The goal is to fit a predictive model to the data as to predict the quality of weightlifting exercise that was performed.

# Getting started
```{r, warning=FALSE}
# Needed packages
library(lattice)
library(ggplot2)
library(caret)
library(randomForest)
library(rpart)

#splitting test and training data
trainingset <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
testingset <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))


```

```{r}
# Removing columns with NA's
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]

# Removing irrelevant data
trainingset   <-trainingset[,-c(1:5)]
testingset <-testingset[,-c(1:5)]

#splitting training data for out-of-sample error:
set.seed(59)
trainsplit <- createDataPartition(y = trainingset$classe, p = 0.70, list = FALSE)
trainingset1 <- trainingset[trainsplit, ]
trainingset2 <- trainingset[-trainsplit, ]
```

# Model 1: Random forest
```{r}
fitControl <- trainControl(method="cv", number=3, verboseIter= FALSE)
fit <- train(classe ~ ., data=trainingset1, method="rf", trControl=fitControl)
pred <- predict(fit, newdata = trainingset2)
confusionMatrix(trainingset2$classe, pred)
```

# Model 2: Decision tree
```{r}
model <- rpart(classe ~ ., data = trainingset1, method = "class")
pred2 <- predict(model, trainingset2, type = "class")
confusionMatrix(trainingset2$classe, pred2)
```
# Selecting a model
Based on both the accuracy and the kappa for both models, the Random Forest model (0.9976, 0.997) does considerably better when compared to the Decision Tree model (0.7519, 0.6842) on the second part of the training set, giving it a smaller ISE. 
This means the model will retrain on the full training dataset using the Random Forest method.

# Retraining
```{r}
fitControl <- trainControl(method="cv", number=3, verboseIter= FALSE)
fit <- train(classe ~ ., data=trainingset, method="rf", trControl=fitControl)
```

# Prediction
```{r}
prediction <- predict(fit, newdata = testingset)
prediction <- as.character(prediction)
prediction
```

### Data source

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13). Stuttgart, Germany: ACM SIGCHI, 2013.